# Non-Communicable Disease Burden (NCD)

```{r, echo=FALSE, message=FALSE}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here, dplyr)

# Import the non-communicable diseases data
data_ncd <- read_csv(here("data", "burden-of-disease-ncd.csv"))
```

## Table of Estimates for NCD Burden Over Time

```{r, echo=FALSE}
# Here render a table for the DALY burden over time for the three countries 
# You should pivot the data to show each country in a separate column.
# Use kable() from the knitr package to render the table.
## Table of Estimates for NCD Burden Over Time

# Filter data for the three specified countries
selected_countries <- data_ncd %>%
  filter(Entity %in% c("Philippines", "Guatemala", "Australia"))

# Pivot the data to show each country in a separate column
pivoted_data <- selected_countries %>%
  select(Entity, Year, `DALYs (Disability-Adjusted Life Years) - Non-communicable diseases - Sex: Both - Age: Age-standardized (Rate)`) %>%
  pivot_wider(names_from = Entity, values_from = `DALYs (Disability-Adjusted Life Years) - Non-communicable diseases - Sex: Both - Age: Age-standardized (Rate)`)

# Render the table using kable()
kable(pivoted_data)
```

## Plot Showing Trends in NCD Burden Over Time 

```{r, echo=FALSE}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.

library(ggplot2)

ggplot(data = selected_countries, aes(x = Year)) +
  geom_line(aes(y = `DALYs (Disability-Adjusted Life Years) - Non-communicable diseases - Sex: Both - Age: Age-standardized (Rate)`, color = Entity)) +
  labs(x = "Year", y = "DALYs", color = "Country") +
  ggtitle("Trend of DALY Burden for Selected Countries") +
  theme_minimal()
```

## Summary of NCD Burden Findings

```{r, echo=FALSE}
# Provide a brief analysis based on the data presented in the table and chart. Highlight any significant findings or patterns. About 3 sentences.
```

In the 30-year period, Australia shows a promising trend of drastic decrease from its values in 1990 to 2016.  Guatemala shows a slow decreasing trend in the DALY burden from 1998 to 2019.  In contrast to the downward trends of DALY burden for Guatemala and Australia, there is an increasing pattern for the Philippines beginning in the year 1991.