# Overall Disease Burden

```{r, echo=FALSE, message=FALSE}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the overall disease data
data_overall <- read_csv(here("data", "burden-of-disease-all-causes.csv"))
```

## Table of Estimates for Overall Disease Burden Over Time

```{r, echo=FALSE}


# Here render a table for the DALY burden over time for the three countries 
# You should pivot the data to show each country in a separate column.
# Use kable() from the knitr package to render the table.

countries <- data_overall %>% 
  select(Entity, DALYs = `DALYs (Disability-Adjusted Life Years) - All causes - Sex: Both - Age: Age-standardized (Rate)`, Year) %>% 
  filter(Entity %in% c("Philippines", "Guatemala", "Australia") ) 

```


```{r, echo=FALSE}
table_countries <- countries %>% 
  pivot_wider(values_from = DALYs, names_from = Entity) 

```


```{r, echo=FALSE}
table_countries_nice <- kable(table_countries)

table_countries_nice
```


## Plot Showing Trends in Overall Disease Burden Over Time


```{r, echo=FALSE}

ggplot(data = countries,
       mapping = aes (x= Year,
                      y= DALYs,
                      group = Entity,
                      color = Entity)) +
  geom_line() +
  labs(title = "Trends in Overall Disease Burden Over Time",
       subtitle = "DALY Burden for selected countries",
       x = "Year",
       y = "DALYs",
       color = "Countries")+
  theme_minimal()


# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
```



## Summary of Overall Disease Burden Findings

We notice an important difference of DALYs between the 3 countries. Guatemala shows much higher DALY burden than Australia and Philippines in 1990, however it is drastically reduced by 2020 reporting similar numbers than Philippines. The three countries show a promising trend over the 30 year period with a decreasing DALY over time.

